Nuclear Wastes: we're kinda doomed

### 2017-01-23 paper.tex
intro done

%\documentclass{article}
\documentclass[preprint,aip,pra]{revtex4-1}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{gensymb}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{physics}
\usepackage{color}
\newcommand{\red}[1]{\textcolor{red}{\bf #1}}

@@

\section{Introduction}
IAEA nuclear waste estimation: over 1000 tons of highly radioactive waste.\cite{iaea08}
Nuclear power estimate:increase by 40\% by 2020. \cite{iaea12}
meep. Mostly focus on nuclear waste from reactors.
Global policy review \cite{r12}

Where they come from:\\
mop

what to do with them:\\
meep

\begin{itemize}
    \item history
    \item nuclear decay mechanism: weak force and tunneling, radioactive decay
    \item reactors?
    \item What is nuclear waste and why is it harmful?
    \item partition and transmutation
    \item Above-ground disposal: look at sites, dry-casks storage
    \item geological disposal ($\star$), Yucca mountain as an example
    \item ocean floor disposal which was done by everyone but not legal any more
    \item transmutation
    \item todo: find sources on the specific health impact of nuclear waste instead of radiation
        in general
\end{itemize}

\section{Radioactivity and nuclear fission}
    Here we detail the nuclear physics underlying most of the things we will talk about
    in this work. We will apply these results in our discussion of nuclear wastes.
    \subsection{Radioactivity}
    \subsubsection{Nuclear shell model vs liquid drop model(?)}
        By solving Schr\"{o}dinger's equation\cite{k88}
    \subsubsection{Binding Energy(?)}

@@

\section{Nuclear waste: classification and properties}
    \red{how much of each is generated and what the impact is; industry vs military}\\
    Different organizations have slightly different classification schemes for nuclear waste.
    However, their definitions of high-level waste (HLW), the most detrimental kind, are relatively consistent.
    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity
    content and the half-life of the radionuclides (see FIG. \ref{fig:scheme}).
    Activity content is a measure of how much radioactivity
    a given amount of waste contains: a form of waste may possesses a high activity content due to
    \red{a high concentration of radionuclides or the presence of radionuclides with high activity.}\cite{iaea09}
    Each form of waste, depending on the half-life, activity content, and temperature, requires different
    disposal methods.
    \begin{figure}[h]
        \centering
        \includegraphics[width=0.7\textwidth]{wastescheme.png}
        \caption{An illustration of the IAEA classification scheme for nuclear waste along with suggested
        disposal methods.\cite{iaea09}}
        \label{fig:scheme}
    \end{figure}
    We will focus on high-level waste (HLW, defined below) in the rest of this work.
    In this section, we will first examine the composition and properties of HLW and then
    comment on other categories of waste and disposal methods.

@@

    \subsection{Transuranic waste(TRU)}
    Transuranic waste is defined as materials that do not belong to high-level waste and are
    contaminated by alpha-emitting radionuclides with
    sufficiently long half-life ($<20$ years) of elements with atomic number $92$ (atomic number of uranium)
    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01}. The activity
    of TRU is usually low but it remains active for long period of time.
    \red{WIPP is only for TRU}

@@

    \section{Spent fuel management}
    Spent fuel from nuclear reactors constitutes the major source of HLW. Spen fuel pools at the
    reactors hold spen fuel cores to cool them down and isolate them for a few years. After that,
    reprocessing can occur to extract the most radioactive plutonium and uranium isotopes and
    convert the waste into more managable form (vitrification). However, reprocessing of commercial
    nuclear spent fueld did not
    happen in this country for various historical reasons. 
    \subsection{Spent fuel pools}
    \red{boron-10 and radiation shielding.}
    The heat output of the spent fuel decreases by $99\%$ after the first year and by another factor of $5$
    by the 5-year mark.\cite{aa12} Therefore, short-term cooling of spent fuel is crucial before
    futher disposal or reprocessing can be done.
    US nuclear power plants generally have an operation cycle of about 2 years. At the end of each operation cycle,
    one-quater to one-third of its fuel assemblies (spent fuel) are removed. The removed spent fuel spend a minimum
    of five years in these water pools to cool.
    Most of the nuclear industry's spent fuel is now stored in spent fuel water pools ``at-reactor (AR)'',
    which means either within the reactor building or in an ajacent spent fuel building linked to the reactor
    AR pools minimize the difficulty of transporting spent fuel.\cite{iaea99}
    US has generated about $65,000$ metric tons of spent fuel, of which $75\%$ is stored in AR pools.\cite{a11}
    by a transfer tunnel. 
    Spent fuel pools  are usually about 15-meter deep water pools, with storage rack at the bottom to hold the fuel assemblies.
    The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
    can provide coooling by circulation. Moreover, water can serve as a shielding from radiation, because it has a
    fairly high density and hence forces the radiated particles (especially the neutrons) to interact with
    the water molecules and lose energy \red{(more precise terms needed, cladding materials)}. Water temperature is maintained
    below 50 $^\circ$C via circulation and cooling. The integrity of the storage can persist well beyond 
    50 years.\cite{a11, iaea99}

@@

    \subsection{Dry cask storage}
    Spent fuel pools have very limited capacities. 
    Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
    After cooling in the spent fuel pools for a few
    years, the spent fuel can be stored in dry casks filled with inert gas. 
    The dry casks are often metal cyliders welded to prevent leakage. Each cask is surrounded by
    additional steel and concrete to provide shielding.
    The heat from the cask is removed by passive
    cooling (air circulation). The Nuclear Regulatory Comission is hoping that the casks will
    last safely for as long as 100 years, yet corrosions and cracking occur in 30 years
    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
    dudring  the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
    \begin{displayquote}
    Leaving the spent fuel onsite for extended periods of time was never intended and is not
    responsible. (Dry cask manufacturer)

    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
    program. (Department of Energy employee speaking as a citizen)
    \end{displayquote}
    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in pools
    that are close or over to their designed limits.\cite{a11,aa12}
    Therefore, there is clearly a need for a long-term disposal plan, which we will epxplore in
    the next section.\\
    \red{(A lot more drama with the NRC going on here, loss of credibility.)}

\section{Disposal of HLW}
    The timeframe of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
    of the waste decreases down to a safe level.
    Deep geological repositories that are stable over timescales naturally offer a good option for disposal.
    It is accepted that geological repositories in a stable environment over 300 meters below ground
    is safter than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}

    Reversible repositories hm.
%    \subsection{Comments on partitioning and transmutation}
    \subsection{Requirements for HLW disposal}
     \subsection{Survey of possibilities}
     transmutation, ocean floor, deep geological disposal
%    \subsubsection{mention ocean disposal and why it was banned('Ndrangheta)}
%    \subsection{Necessity of geological disposal sites}
%    \subsection{Requirements of disposal from physics}
     \subsection{Deep borehole disposal}
     \subsection{Deep geological repository}
%
\section{Deep Geological Repository}
    Therefore, long-term storage before final disposal is crucial. 
    \subsection{something general}
    \subsection{Yucca Mountain}
    \subsection{Something related to the French Ciegeo project: licensed reversible disposal site}
    \subsection{Risk calcs: geophysics, mention WIPP}

>>>>

%\documentclass{article}
\documentclass[nofootinbib,preprint,aip,pra]{revtex4-1}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{gensymb}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{physics}
\usepackage{color}
\newcommand{\red}[1]{\textcolor{red}{\bf #1}}

@@

\section{Introduction}
Nuclear power continues to play an important role in global energy supply.
The International Atomic Energy Agency (IAEA) estimated that the power generated by nuclear power
is continuing to grow and will increase by $40\%$ by 2030.\cite{iaea12}
Whereas the ``front end'' of the nuclear fuel cycle, for example reactor physics and power output, draws
the most attention from the public, the ``back end'' of the fuel cycle, for example waste management and
disposal, deserves equal, if not more, attention.
Currently, the IAEA estimates that there are over 10,000 metric tons of highly radioactive nuclear waste 
generated worldwide annually,
most of which is spent nuclear fuel from nuclear reactors.\cite{iaea08, r12}
The world will have over $373,800$ metric tons of highly radioactive waste stock by the end of 2020.\cite{r12}
Highly radioactive waste remains hazardous for up to $10^6$ years after it is produced and thus requires extremely
careful 
Nevertheless, there is currently no permanent repository in the world to dispose of these highly radioactive
waste.~\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
    waste repository, but only
for waste classified as Transuranic, which, by the Nuclear Regulatory Commission standard, does not have particularly
high radioactivity but only long half-life (See Sec.~\ref{sec:tru}). In particular, WIPP does not accept spent nuclear
fuel.}  More than $75\%$ of the spent nuclear fuel worldwide is stored onsite at the nuclear power plants in
water pools that are close to their limiting capacity.\cite{aa12}

Two weeks after the davastating earthquake and tsunami hit the Fukushima Daiichi Nuclear Power Plant in Japan,
the water pool cooling system was cut off from power and officials feared that
the highly radioactive waste in the pools would catch fire and send radioactive smoke across a much larger part of Japan, including Tokyo.
The Nuclear Regulatory Comission (NRC), the independent government agency of the U.S. government overseeing 
nuclear energy safety,
estimates that a major fire in a U.S. spent fuel pool is expected to
displace 3.4 million people in an area larger than New Jersy.\cite{s16}

As we can see, nuclear waste management and disposal is becoming a increasingly urgent matter globally.
In light of the high stake,
the uncertainty of the stability of a geological repository over a extremely long timescale has evolved into 
a topic not only of scientific contention but also of policy debate.
How well is the predicative power of science? How much trust should we put into the uncertainty estimates
(i.e. how well can science predict its predicative power)?
In this work, we will attempt to examine some of the scientific issues underlying nuclear waste management
and disposal. We will start by building up some
language and knowledge necessary to understand how nuclear waste is generated (Section~\ref{sec:phys}),
then apply those principles to study
what physical properties nuclear waste possesses (Section~\ref{sec:waste}), its management and temporary storage
(Section~\ref{sec:temp}), and its disposal (Section~\ref{sec:disposal}).

\section{Radioactivity and nuclear fission}
\label{sec:phys}
    Here we detail the nuclear physics underlying most of the things we will talk about
    in this work. We will apply these results in our discussion of nuclear wastes.
    \subsection{Radioactivity}
    \subsubsection{Nuclear shell model vs liquid drop model(?)}
        By solving Schr\"{o}dinger's equation\cite{k88}
    \subsubsection{Binding Energy(?)}

@@

\section{Nuclear waste: classification and properties}
\label{sec:waste}
    \red{how much of each is generated and what the impact is; industry vs military}\\
    Different organizations have slightly different classification schemes for nuclear waste.
    However, their definitions of high-level waste (HLW), the most detrimental kind, are relatively consistent.
    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity
    content and the half-life of the radionuclides (see FIG. \ref{fig:scheme}).
    Activity content is a measure of how much radioactivity
    a given amount of waste contains: a form of waste may possesses a high activity content due to
    \red{a high concentration of radionuclides or the presence of radionuclides with high activity.}\cite{iaea09}
    Each form of waste, depending on the half-life, activity content, and temperature, requires different
    disposal methods.
    \begin{figure}[h]
        \centering
        \includegraphics[width=0.7\textwidth]{wastescheme.png}
        \caption{An illustration of the IAEA classification scheme for nuclear waste along with suggested
        disposal methods.\cite{iaea09}}
        \label{fig:scheme}
    \end{figure}
    We will focus on high-level waste (HLW, defined below) in the rest of this work.
    In this section, we will first examine the composition and properties of HLW and then
    comment on other categories of waste and disposal methods.

@@

    \subsection{Transuranic waste(TRU)}
    \label{sec:tru}
    Transuranic waste is defined as materials that do not belong to high-level waste and are
    contaminated by alpha-emitting radionuclides with
    sufficiently long half-life ($<20$ years) of elements with atomic number $92$ (atomic number of uranium)
    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01}. The activity
    of TRU is usually low but it remains active for long period of time.
    \red{WIPP is only for TRU}

@@

    \section{Spent fuel management}
    \label{sec:temp}
    Spent fuel from nuclear reactors constitutes the major source of HLW. Spen fuel pools at the
    reactors hold spen fuel cores to cool them down and isolate them for a few years. After that,
    reprocessing can occur to extract the most radioactive plutonium and uranium isotopes and
    convert the waste into more managable form (vitrification). However, reprocessing of commercial
    nuclear spent fueld did not
    happen in this country for various historical reasons. 
    \subsection{Spent fuel pools}
    \red{boron-10 and radiation shielding.}
    The heat output of the spent fuel decreases by $99\%$ after the first year and by another factor of $5$
    by the 5-year mark.\cite{aa12} Therefore, short-term cooling of spent fuel is crucial before
    futher disposal or reprocessing can be done.
    US nuclear power plants generally have an operation cycle of about 2 years. At the end of each operation cycle,
    one-quater to one-third of its fuel assemblies (spent fuel) are removed.
    Most of the nuclear industry's spent fuel is now stored in spent fuel water pools ``at-reactor (AR)'',
    which means either within the reactor building or in an ajacent spent fuel building linked to the reactor
    AR pools minimize the difficulty of transporting spent fuel.\cite{iaea99}
    US has generated about $65,000$ metric tons of spent fuel, of which $75\%$ is stored in AR pools.\cite{a11}
    by a transfer tunnel. 
    Spent fuel pools  are usually about 15-meter deep borated water pools, with storage rack at the bottom to hold the fuel assemblies.
    The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
    can provide coooling by circulation. Moreover, water can serve as a shielding from radiation, because it has a
    fairly high density and hence forces the radiated particles (especially the neutrons) to interact with
    the water molecules and lose energy \red{(more precise terms needed, cladding materials)}. Water temperature is maintained
    below 50 $^\circ$C via circulation and cooling. The integrity of the storage can persist well beyond 
    50 years.\cite{a11, iaea99}

@@

    \subsection{Dry cask storage}
    Spent fuel pools have very limited capacities. 
    Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
    The dry casks are often metal cyliders welded to prevent leakage. Each cask is surrounded by
    additional steel and concrete to provide shielding.
    The heat from the cask is removed by passive
    cooling (air circulation). The Nuclear Regulatory Comission is hoping that the casks will
    last safely for as long as 100 years, yet corrosions and cracking occur in 30 years
    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
    dudring  the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
    \begin{displayquote}
    Leaving the spent fuel onsite for extended periods of time was never intended and is not
    responsible. (Dry cask manufacturer)

    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
    program. (Department of Energy employee speaking as a citizen)
    \end{displayquote}
    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
    packed pools
    that are close to or over their designed limits.\cite{a11,aa12}
    Therefore, there is clearly a need for a long-term disposal plan, which we will epxplore in
    the next section.\\
    \red{(A lot more drama with the NRC going on here, loss of credibility.)}

\section{Disposal of HLW}
\label{sec:disposal}
    The timeframe of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
    of the waste is predicted to decline to a safe level.
    Deep geological repositories that are stable over timescales naturally offer a good option for disposal.
    It is accepted that geological repositories in a stable environment over 300 meters below ground
    is safter than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
    In this section, we first state the requirements for a HLW geological disposal solution, and then
    we look at  deep geological repository, and its alternatives (or variations): deep boreholes storage
    and reversible repository.
    \subsection{Requirements for HLW disposal}
    \subsection{Survey of possibilities}
    transmutation, ocean floor, deep geological disposal
    \subsection{Deep borehole disposal}
        \subsection{Deep geological repository: Yucca Mountain}
        \subsubsection{Risk}
    \subsection{Reversible disposal sites}
